Enabling the Use of GPU Virtualization in Cloud Environments

### 2015-10-08 paper.tex
rev sergio

\abstract{The use of accelerators in order to reduce the execution time of some 
application has become more popular during last years. These devices increment 
the computational power of a node thanks to its parallel architecture.
This trend has led several cloud vendors such as Amazon Web Services or some 
middlewares such as OpenStack to add to their current facilities, or to allow, 
instances of virtual machines equipped with NVIDIA GPUs. 
Accomplishing this requirements needs that the guest hosts must be equipped with 
GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
The solution presented in this work is based on GPU virtualization and shareability 
in order to reach an equilibrium between supply and demand of accelerators from the users. 
Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
With this software configuration, the GPUs can be accessed from any VM getting rid 
of the necessity of having physical GPUs in the guest host. 
Moreover, our project studies the viability of these join by using a public cloud 
service configuration and develops a module for OpenStack middleware in order to 
add the needed support for the virtualized devices and the logic to manage them. 
The results show a feasible configuration which adds flexibility to the current and well-known 
cloud solutions. 
}

@@

@@

\subsubsection{Network behaviour}
As it is shown in the Table~\ref{table:awsInstances}, each instance has different 
interconnection and taking into account that the use of the network 
is critical when GPU virtualization is applied, a simple test to verify the 
real network bandwidth is needed.

@@

The IPERF\footnote{\url{http://iperf.fr/}} tool has been executed between the instances showed in 
the Table~\ref{table:iperf} in order to know the real bandwidth.
Therefore, we can assure that the nomenclature ``High'' used in the 
network column correspond to a 1~Gb interconnection and the 10~Gb has 
a real bandwidth of 7.5~Gb/s.
Moreover, it looks like the instances equipped with a ``High'' interconnection 
are software-limited up to 1~Gb/s due to the theoretical and real bandwidth 
match perfectly. The gap between real and theoretical can be observed with 
the 10~Gb interconnection which reaches up to 7.5~Gb/s.

\subsubsection{GPUs}
As it has been said before, an instance is based in a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}. 
As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is a real GPU. 
\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Name &  Data Movement & Network & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

\subsection{Testbed Scenarios}
In this section, the different client-server configurations 
are explained. All the scenarios are based in the maximum number of instances 
that a user can freely select without submitting a formal request. 
In particular, can be launched 5 of ``g2.2xlarge'' and 2 of ``g2.8xlarge''.

@@

\begin{itemize}
\item Scenario A (Figure~\ref{subfig:aws1}) shows the most common 
and used hardware configuration in real clusters, where each node is 
populated with one GPU. Here, a simple node is able to use up to 5 GPUs 
using the ``High'' interconnection.

@@

Once the scenarios are configured from the point of view of hardware, 
the {rCUDA} software needs to be installed in order to add 
the required flexibility to the system. So, the {rCUDA} server has been 
executed in the server nodes and the {rCUDA} libraries have been installed 
in the node that is acting as a client.

@@

The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, an application purely computational. 
It was launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
We have removed the results of the Scenario~\ref{subfig:aws2} for clarity ought to they are exactly the same that the obtained from Scenario~\ref{subfig:aws3} up to 8 GPUs. 
In other words, we could state that the results until the 8th GPU would correspond to the Scenario~\ref{subfig:aws2} and thereafter to the Scenario~\ref{subfig:aws3}.
Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario~\ref{subfig:aws1} could increase the throughput because
unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
% \begin{figure}[!t]
%   \centering
%   \includegraphics[width=\linewidth]{images/aws-mont2.png}
%   \caption{Montecarlo execution time}
%   \label{fig:mont-exe}
% \end{figure}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{images/mont.pdf}
  \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local and remote GPUs.}
  \label{fig:mont-opt}
\end{figure}

The other application, LAMMPS\footnote{\url{http://lammps.sandia.gov}}, is a classical molecular dynamics simulator that can be applied at the atomic, meso, or continuum scale. 
From the implementation perspective, this multi-process application employs at least one GPU to host its processes, but can benefit from the use of multiple GPUs.
The Subfigure~\ref{subfig:lammps1} shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
Furthermore, for the execution on remote GPUs, can be seen a little difference between both networks, the ``High'' network performs worse than ``10 Gb'' network.
Thus, if we compare the results of a LAMMPS execution with a larger problem (see Subfigure~\ref{subfig:lammps2}), although CUDA continues performing better, the interesting point is the execution time on remote GPUs. 
They are almost the same even with different networks what means that the transfers are making the interconnection network the bottleneck, and enlarging the problem size
pays off the drawback of a slower network.

@@

The last conclusion we drew was related to GPU-shareability. 
As we deduced, AWS has GPU-capable physical machines reserved which could be idle until a GPU-instance request were made.
To have such an expensive accelerators in a standby state is counter-productive. So, it would make more sense if we had less devices, accessible from any machine, though.

@@

%\section{\uppercase{Discussion}}
%\label{sec:discussion}
%-Volver a hacer un resumen

>>>>

\abstract{The use of accelerators, such as GPUs, to reduce the execution time of GPGPU applications has become more popular during last years. 
These devices increment the computational power of a node thanks to its parallel architecture.
This trend has led several cloud vendors such as Amazon Web Services or some middlewares such as OpenStack to add to their current facilities, or to allow, 
instances of virtual machines equipped with NVIDIA GPUs. 
Accomplishing these needs, the guest hosts must be equipped with GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between supply and demand of accelerators. 
Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
With this software configuration, the GPUs can be accessed from any VM getting rid 
of the necessity of having physical GPUs in the guest host. 
Moreover, our project studies the viability of these join by using a public cloud 
service configuration and develops a module for OpenStack middleware in order to 
add the needed support for the virtualized devices and the logic to manage them. 
The results show a feasible configuration which adds flexibility to the current and well-known 
cloud solutions. 
}

@@

****** FALTA UNIR AMBAS PARTES DEL TRABAJO ******

@@

\subsubsection{Networking}
As it is shown in the Table~\ref{table:awsInstances}, each instance has a different 
network and taking into account that the use of the bandwidth 
is critical when GPU virtualization is applied, a simple test to verify the 
real network bandwidth is needed.

@@

The IPERF\footnote{\url{http://iperf.fr/}} tool has been executed between the instances showed in 
the Table~\ref{table:iperf} in order to know the real bandwidth.
Therefore, we can assure that the nomenclature ``High'' used in the 
network column correspond to a 1~Gb interconnection network and the 10~Gb has 
a real bandwidth of 7.5~Gb/s.
Moreover, it looks like the instances equipped with a ``High'' interconnection network
are software-limited up to 1~Gb/s due to the theoretical and real bandwidth 
match perfectly. The gap between real and theoretical can be observed with 
the 10~Gb interconnection which reaches up to 7.5~Gb/s.

\subsubsection{GPUs}
As it has been said before, an instance is based on a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}. 
As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is an ad-hoc GPU. 
\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Name &  Data Movement & Network & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

\subsection{Testbed Scenarios}
In this section, the different client-server configurations 
are explained. All the scenarios are based on the maximum number of instances 
that a user can freely select without submitting a formal request. 
In particular, the maximum number of ``g2.2xlarge'' is 5 and 2 for the ``g2.8xlarge'' instances.

@@

\begin{itemize}
\item Scenario A (Figure~\ref{subfig:aws1}) shows the most common 
and used hardware configuration in real clusters, where each node is 
populated with one GPU. Here, a simple node is able to use up to 5 GPUs 
using the ``High'' network.

@@

Once the scenarios are configured from the point of view of hardware, 
the {rCUDA} software needs to be installed in order to add 
the required flexibility to the system. So, the {rCUDA} server is 
executed in the server nodes and the {rCUDA} libraries are installed 
in the node that is acting as a client.

@@

The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, an application purely computational. 
It was launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
We have removed the results of the Scenario B for clarity ought to they are exactly the same that the obtained from Scenario C up to 8 GPUs. 
In other words, we could state that the results until the 8th GPU would correspond to the Scenario B and thereafter to the Scenario C.
Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario A could increase the throughput because
unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
% \begin{figure}[!t]
%   \centering
%   \includegraphics[width=\linewidth]{images/aws-mont2.png}
%   \caption{Montecarlo execution time}
%   \label{fig:mont-exe}
% \end{figure}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{images/mont.pdf}
  \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local and remote GPUs.}
  \label{fig:mont-opt}
\end{figure}

The other application, LAMMPS\footnote{\url{http://lammps.sandia.gov}}, is a classical molecular dynamics simulator that can be applied at the atomic, meso, or continuum scale. 
From the implementation perspective, this multi-process application employs at least one GPU to host its processes, but can benefit from the use of multiple GPUs.
The Subfigure~\ref{subfig:lammps1} shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
Furthermore, for the execution on remote GPUs, a little difference between both networks can be appreciated, so, the ``High'' network performs worse than ``10 Gb'' network.
Thus, if we compare the results of a LAMMPS execution with a larger problem (see Subfigure~\ref{subfig:lammps2}), although CUDA continues performing better, the interesting point is the execution time on remote GPUs. 
They are almost the same even with different networks what means that the transfers are making the interconnection network the bottleneck, and enlarging the problem size
pays off the drawback of a slower network.

@@

The last conclusion we will draw is related to GPU-shareability. 
As we deduced, AWS has GPU-capable physical machines reserved which could be idle until a GPU-instance request were made.
To have such an expensive accelerators in a standby state is counter-productive. So, it would make more sense if we had less devices, accessible from any machine, though.

@@

\subsection{Discussion}
\label{sec:discussion}
%-Volver a hacer un resumen
